---
title: "Douglas color cards"
author: "Nicolas Hahn, Eric Jonas, Pavlo Kravets, Simon Schäfer"
date: "`r Sys.Date()`"
output: bookdown::gitbook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/setup.R")
```

# Douglas color cards

The Douglas color cards are used to determine someones skin color to find the most appropriate cosmetics products - e.g. the perfect foundation. They show 64 squares of various colors. The center four squares get punched out.

```{r colorcard, echo=FALSE, warning=FALSE}
display_color_card(master_colors, "Crow", "Ccol", "Lab", 2)
```

The color cards are placed in front of ones facial skin and a photo is taken. The visual information about the skin color (visible at the four center squares) and the reference color squares is evaluated by an app and the closest skin color type is calculated.

Each sheet contains 42 color cards, each with a QR code. The QR codes allow the app to identify the precise location of each color card on the sheet. Below is a visualization of the first sheet in our dataset.

```{r colorcard first sheet, echo=FALSE, warning=FALSE}
display_color_sheet(lab_colors, color_sheet_idx = 1)
```

## Quality assurance

To assure the best skin color typification it is crucial to assure a consistent production of the color cards. Especially the reference color squares colors variance should be as low as possible.

The existence of QR codes indicating the position on a sheet reduces the requirements to meet because thus only the variance of reference color squares at the same position of the sheets have to be as low as possible. Differences between reference color squares among different positions on a sheet can be taken care of in the app once the positional information is evaluated from the QR code.

## Data description

The data given consists of two files. One file holds the master color information given in the Lab colorspace as well as in the CMYKS color space. In the latter S stands for special and might be some kind of skin color pigment.

The second file holds color information obtained from 13 printed sheets. Thus we find the Lab color information for all 64 squares for 42 color cards for 13 sheets resulting in `r format(64*42*13*3, scientific=FALSE)` numeric values.

## Objectives

* Analyse where variance or difference in color is noteworthy high / low
* Visualize variance and differences in colors
* Use the CMYKS color information to match the Lab color values

### Questions (formulated before exploration)

1. Is the color difference greater within a sheet (among positions) or across sheets?  (Simon Schäfer: R-Code ✔)
1. Is there a difference in dE regarding skin or non skin color / calibration color?  (Pavlov)
1. How is the color special used (regression)? (Nico)
1. What dE value can be seen on our screen? (Simon Schäfer: R-Code ✔, Text ✔)
1. Color difference of mean over all cards to master target: is it in some uncertainty area? (Eric Jonas: R-Code ✔, Text ✔)
1. Are some Lab values more often in master than in mean? (Eric Jonas: R-Code ✔, Text ✔)
1. How close are CMYK and Lab values? (Eric Jonas: R-Code ✔, Text ✔)
1. Is there a correlation between dE(CMYK colors, Lab colors) and S? (Eric Jonas: R-Code ✔, Text ✔))
1. Are there colors that are very far off the master target? <- Field 18 (Why; see Lab values)
1. Are findings on L, a or b also there in dE? (On the subjective meaning)?
1. Are there especially good / bad cards? (Consistently / target specific) (Pavlo)

## Color spaces (Pavlo)

Lab, CMYK, rgb, CMYKS?

### Analysis of S in CMYKS

How does the S value, representing skin tone, influence CMYK values? When comparing the Lab Master with the CMYK Master excluding S, a notable difference emerges, suggesting that the S value significantly impacts CMYK values when converted to Lab.

```{r CMYK and Lab master colorcard, echo=FALSE, warning=FALSE}
grid.arrange(display_color_card(master_colors, "Crow", "Ccol", "CMYK", 2, 12), 
             display_color_card(master_colors, "Crow", "Ccol", "Lab", 2, 12), 
             ncol = 2)
```

A detailed examination is depicted in the graphic below:

```{r CMYK and Lab master colorcard difference, echo=FALSE, warning=FALSE, fig.width=8, fig.height=8}
master_colors_cmyk <- master_colors %>%
  rowwise() %>%
  mutate(L = cmyk_to_lab(C, M, Y, K)[1],
         a = cmyk_to_lab(C, M, Y, K)[2],
         b = cmyk_to_lab(C, M, Y, K)[3]) %>%
  as.data.frame()
master_colors_cmyk_differences <- generate_color_difference_df(master_colors, master_colors_cmyk)
plot_card_differences_to_master(master_colors_cmyk_differences, master_colors)
```

Given the ambiguity in the CMYK-to-Lab conversion method, this chapter explores correlations between Lab values and S values, employing machine learning models to enhance our understanding of S values and their role in CMYK conversion.

**TODO: ADD CONCLUSION (Nico)**

#### Analysis of Lab Correlations with S

In this section, we investigate the relationships between S values and Lab values by correlation.

Examining S values reveals that brownish colors tend to have higher S values. The left side of the gradient—from black to white—shows an S value of 0. This observation suggests that luminance has minimal, if any, influence on S values.
The tile plots below visualizes this relationship. It illustrates how S values vary across different color points in the dataset.

```{r S tile plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
ggplot(master_colors, aes(x = Ccol, y = Crow)) +
  geom_tile(aes(fill = S), color = "white") +
  geom_text(aes(label = S), color = "black") +
  scale_fill_gradient(low = "white", high = "#4682B4", name = "S-value") +
  labs(title = "Tile Plot of S-values",
       x = "Column",
       y = "Row") +
  theme_minimal()
```

Next, we examine the correlations between individual Lab dimensions (L, a, b, Chroma) and S.

```{r Lab correlations with S, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=8}
grid.arrange(plot_correlation_with_categories(master_colors, "L", "S"),
             plot_correlation_with_categories(master_colors, "a", "S"),
             plot_correlation_with_categories(master_colors, "b", "S"),
             plot_correlation_with_categories(master_colors, "Chroma", "S"),
             ncol = 2, nrow = 2)
```

The L dimension appears to correlate strongly with S values, when excluding cases where S equals 0. However, this correlation is misleading as our analysis indicates that L values do not significantly influence S values.

Notably, the S values also show strong correlations with 'a' and 'b' and the strongest correlation with Chroma. Chroma in the Lab color space quantifies the saturation or colorfulness of a color. It measures how far a color is from being neutral (gray). Chroma in the Lab color space is calculated using the formula:

\[ C = \sqrt{a^2 + b^2} \]

where 'a' and 'b' represent the color opponent dimensions in the Lab color model.
Pearson correlations tend to be higher than Spearman rank correlations, implying a predominantly linear relationship between S and Lab values. However, outliers significantly influence Pearson coefficients, challenging the assumption of a straightforward linear dependency between S and Lab values across all data points. Moreover, upon examining the scatter plots, a strict linear relationship is not evident.

From this analysis, we conclude that the color values 'a' and 'b' exert a stronger influence on S than luminance.

#### Analysis of relationship with machine learning models

**TODO (Nico)**

Also see 1.6.

## Color difference

One can analyse difference in color in at least three ways:

1. Qualitative: Looking at them - but this takes a while if thee are many colors
1. Quantitative: Compare the numeric values of each color space dimension
1. Quantitative: Compare all color space dimensions at once with a proper function (i. e. \( \Delta \) E)

### Comparing color space dimensions one by one

In this chapter, we will compare the colorspaces with the master based on their different dimensions. 

To compare the color spaces, we generated a mean color card derived from all color cards in our *Lab_Measurement* dataset. 
This mean color card serves as a baseline against the reference data of the master. Below is a representation of the mean color card.
```{r mean colorcard, echo=FALSE, warning=FALSE}
display_color_card(mean_lab_color_card, "Row", "Col", "Lab", 2)
```

#### Analysis of field discrepancies

```{r colorcard mean vs master, echo=FALSE, warning=FALSE, fig.width=8, fig.height=10}
plot_card_vs_master(mean_lab_color_card, master_colors)
```

Our examination reveals notable fluctuations across various fields. Particularly, the 18th field exhibits significant deviations from the master's data. Additionally, a trend is observed where decreases in 'a' and 'b' values coincide with increases in 'L' values, notably in fields 54 to 56.
Overall, it is evident that the mean of the color cards approximately matches the master in most other fields.

#### Analysis of density distributions

Examining the distribution of values across individual channels provides insights into the predominant brightness and color characteristics. The following graphic displays the median density of all cards along with the 95% confidence interval.

```{r colorcard distributions, echo=FALSE, fig.height=6, warning=FALSE, fig.align='center'}
plots <- plot_density_vs_master(lab_colors_master_shape, master_colors)
grid.arrange(plots[["L"]], plots[["a"]], plots[["b"]], nrow = 3)
```

The distributions reveal distinct characteristics across the channels: 

- **L dimension:** Notably, there is a leftward shift in values compared to the master, indicating that the dataset tends to have higher brightness levels than the master reference. Additionally, the L channel shows significantly higher variability compared to the other dimensions, where variability is notably low.
- **a dimension:** There is a noticeable deviation from the master particularly around 'a' values near zero, suggesting limitations in accurately representing subtle color variations within this range.
- **b dimension:** Similar to 'a', deviations are observed especially around 'b' values between -10 and 40, indicating representations that differ from the master inside this range.

### Comparing colors with distance function

A function to compare two colors given in Lab color space is \( \Delta \) E - also called dE in this report. Actually there are multiple versions of [Delta E](https://zschuessler.github.io/DeltaE/learn/). The first was given 1976 and just calculates the euclidean distance in the Lab color space:

\[\Delta E_{ab}=\sqrt{(L_2-L_1)^2+(a_2-a_1)^2+(b_2-b_1)^2}\]

In this report we use the most recent \( \Delta \) E formula from 2000. It is much more complex but solves problems when calculating color difference for colors with high lightness value.

The aim of dE is that the calculated distance value can be interpreted in a meaningful way for humans. The following table can be a taken as a guide but there exceptions are possible.

| Delta E   |	Perception                                |
|-----------|-------------------------------------------|
| <= 1.0    | Not perceptible by human eyes.            |
| 1 - 2     | Perceptible through close observation.    |
| 2 - 10    | Perceptible at a glance.                  |
| 11 - 49   | Colors are more similar than opposite     |
| 100       |	Colors are exact opposite                 |

#### Color difference on a single card

Comparing the first color cards colors with the master colors shows that there is one color that is pretty far from the master color (3rd row, 2nd column). There are some more fields where we spotted a difference at first glimpse:

* lilac: 8th row, 5th column
* blue: 8th row, 7th column
* dark grey: 7th row, 1st column

The visualization used to compare the colors shows the 64 colors of the color cards according to their position on the color card. In each tile there are two colored circles which show the master color (left circle) and the sample color (right circle). The background color of the tiles corresponds to the dE-value. Light blue indicates a high dE-value. Dark blue ones a low dE-value.

```{r, echo=FALSE}
color_compar_tiles <- function(spotsize, spotdistance, tilesize = 1, tilehighlightcolor = "#000000", lwd = 1) {
  color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow),
            lwd = lwd, width=tilesize, height=tilesize) +
  scale_color_manual(values = c("#FFFFFF00", tilehighlightcolor)) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+spotdistance, y = Crow, color = Color), size = spotsize) +
  scale_color_identity() +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-spotdistance, y = Crow, color = Color), size = spotsize
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
}
```

```{r dE_colorcard, echo=FALSE, fig.width=8, fig.height=8}
color_compar_tiles(spotsize = 11, spotdistance = 0.25)
```

To spot the color difference it is very helpful to let the colors to compare overlap. Thus we can find more colors that are distinguishable by us on our screen. Almost all grey colors are separable from their master color. So are most of the brown colors. Most of the other colors seem to be equal. An exception is the red color (3rd row, 8th column).


```{r dE_colorcard_2, echo=FALSE, fig.width=8, fig.height=8}
color_compar_tiles(spotsize = 13, spotdistance = 0.175)
```

In order to check if the cut off value of 2 holds for our example we show the same picture again but highlight which colors should be distinguishable in theory. It tells us that we sould be able to spot the difference for the following colors but we did not:

* magenta: 4th row, 8th column
* olive: 1st row, 2nd column
* lavender: 1st row, 5th column

```{r dE_colorcard_3, echo=FALSE, fig.width=8, fig.height=8}
color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference > 2),
            lwd = 1.5, width=0.95, height=0.95) +
  scale_color_manual(values = c("#FFFFFF00", "#CC3333")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = Color), size = 13) +
  scale_color_identity() +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 13
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
```

Finally we have a look at which color pairs should be no visible difference. Here we found no exception to the rule of thumb given.

```{r dE_colorcard_4, echo=FALSE, fig.width=8, fig.height=8}
color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference < 1),
            lwd = 1.5, width=0.95, height=0.95) +
  scale_color_manual(values = c("#FFFFFF00", "#88CC00")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = Color), size = 13) +
  scale_color_identity() +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 13
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
```

#### Correlation \( \Delta \) E with Lab Dimensions

In the previous chapter, we analyzed the difference of \( \Delta \) E across various color fields. This chapter explores the influence of channels on \( \Delta \) E.

```{r Lab correlations with deltaE, echo=FALSE, warning=FALSE, message=FALSE}
grid.arrange(plot_correlation(mean_lab_color_differences, "L", "Difference"),
             plot_correlation(mean_lab_color_differences, "a", "Difference"),
             plot_correlation(mean_lab_color_differences, "b", "Difference"),
             nrow = 3)
```

From the scatter plots, it is evident that there is generally no or very weak linear correlation between \( \Delta \) E and the Lab dimensions. Interestingly, the Spearman coefficient highlights a notable correlation between L and \( \Delta \) E values. This suggests that the perception of color differences may not be solely tied to the individual color channels a and b in a straightforward linear manner. Rather, factors influencing the perception of lightness differences appear to be more intricate and multifaceted. These findings prompt further exploration into the nuanced aspects of color perception beyond basic channel influences.

#### Color variance among cards

```{r, echo=FALSE}
DeltaE_map_for_sheet <- function(data) {
  data %>% 
    ggplot() +
    geom_tile(aes(fill = Difference, x = Ccol, y = Crow)) +
    coord_fixed() +
    theme(
      axis.text=element_text(size=12),
      axis.title=element_text(size=14,face="bold"),
      plot.caption = element_text(size=12, hjust = 0)
      )
}
```

In order to get a feeling if the color difference compared to the masters colors varies more or less regarding to the color cards position on the sheet or among the sheets at the same position one can have a look at small multiples for both cases. Displaying all color cards of all sheets is possible on a huge screen but not in this report. Anyhow - as one can see - comparing the separated dE tile maps visually is not easy.

Looking at all color cards in a specific position among all 13 sheets shows little difference. Looking at all 42 color cards of a single sheet shows that there might be two slight patterns. One might find a) a light diagonal from bottom left to top right (see cards in first row) or b) a dark diagonal from top left to bottom right (see cards in last row). We will investigate this further with other visualization techniques.

```{r dE_colorcards_topleft_all_sheets, echo=FALSE, fig.width=8, fig.height=8}
DeltaE_map_for_sheet(color_differences %>% filter(Row == 1, Column == 1)) +
  labs(caption = "dE for all color cards at the top left position of all sheets") +
  facet_wrap(Sheet ~.) +
  theme(
    axis.title = element_blank()
  )
```

```{r dE_colorcards_1st_sheet, echo=FALSE, fig.width=8, fig.height=8}
DeltaE_map_for_sheet(color_differences %>% filter(Sheet == 1)) +
  labs(caption = "dE for all color cards of the first sheet") +
  facet_grid(Row ~ Column) +
  xlab("Column on sheet") + ylab("Row on sheet")
```

```{r, echo=FALSE, echo=FALSE, fig.width=8, fig.height=12}
color_dispersion_on_sheet1 <- color_differences %>%
  filter(Sheet == 1) %>% 
  mutate(Aggregated_by = "Sheet")

color_dispersion_on_target1_1 <- color_differences %>%
  filter(Row == 1, Column == 1) %>% 
  mutate(Aggregated_by = "Target")

color_dispersion_on_sheet1 %>% bind_rows(
  color_dispersion_on_target1_1
) %>% filter(Field != 18) %>% 
  ggplot() +
  geom_boxplot(aes(x = Aggregated_by, y = Difference)) +
  facet_grid(Crow ~ Ccol, switch = "both", as.table=FALSE) +
  coord_cartesian(ylim = c(0, 6.5))
```

```{r, echo=FALSE, message=FALSE}
color_dispersion_over_sheets <- color_differences %>% group_by(Sheet, Field) %>% 
  summarise(median_dE = median(Difference), count = n(), mad_dE = mad(Difference)) %>% 
  mutate(Aggregated_by = "Sheet") %>% 
  ungroup()

color_dispersion_over_targets <- color_differences %>% group_by(Field, Row, Column) %>% 
  summarise(median_dE = median(Difference), count = n(), mad_dE = mad(Difference)) %>% 
  mutate(Aggregated_by = "Target") %>% 
  ungroup()

color_dispersion_compare <- color_dispersion_over_sheets %>% bind_rows(
  color_dispersion_over_targets
)
```

**Color variety in general lower within a target position than over a sheet. Makes sense to generate QR code to inform app which color a printed square is actually closer to (instead of master color).**

```{r, echo=FALSE, echo=FALSE, fig.width=8, fig.height=12}
color_dispersion_compare %>% left_join(
  color_differences %>% select(Field, Crow, Ccol) %>% unique(),
  by = join_by(Field)
) %>% #filter(Field != 18) %>% 
  ggplot() +
  geom_boxplot(aes(x = Aggregated_by, y = mad_dE), alpha = 0.5) +
  facet_grid(Crow ~ Ccol, as.table = FALSE, switch = "both")
```

**No real difference in median color distances. color distances more spread for targets. (Less individual values per group for target 13 vs 42 over sheets and more groups could explain wider box and whisker)**

```{r, echo=FALSE, echo=FALSE, fig.width=8, fig.height=12}
color_dispersion_compare %>% left_join(
  color_differences %>% select(Field, Crow, Ccol) %>% unique(),
  by = join_by(Field)
) %>% filter(Field != 18) %>% 
  ggplot() +
  geom_boxplot(aes(x = Aggregated_by, y = median_dE), alpha = 0.5) +
  facet_grid(Crow ~ Ccol, as.table = FALSE, switch = "both")
```

## New Reference colors by targets

**Colors vary less among targets. What is the the "new" reference color the colors are scattered around? Compare to master color for one target.**

## Douglas CMYKS-Formula

The Douglas master color data yields colors in a unique CMYKS space. Here S stands for special. Ignoring this value is a bad idea if one wants to match the reference color. Converting just the CMYK values leads to:

```{r colors_only_CMYK, echo=FALSE, fig.width=8, fig.height=8}
mastercolor_cmyk_difference <- master_colors %>% rowwise() %>% mutate(rgb_by_cmyk = cmyk_to_rgb(C,M,Y,K)) %>% 
  bind_cols(
    map_dfr(srgb_to_Lab, .x = .$rgb_by_cmyk) 
  ) %>% bind_cols(
    dE(.[c("L", "a", "b")], .[c("Lab.L", "Lab.a", "Lab.b")]) %>% as.data.frame() %>% setNames("Difference")
  )

mastercolor_cmyk_difference %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference > 2),
            lwd = 1.5, linetype = 1, width=0.9, height=0.9) +
  scale_color_manual(values = c("#FFFFFF00", "#CC3333")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = rgb_by_cmyk), size = 14) +
  scale_color_identity() +
  geom_point(
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 14
  ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
  )
```

```{r, echo=FALSE}
options(scipen = 1, digits = 2)

color_regression_linear <- lm(data = master_colors, cbind(L, a, b) ~ C + M + Y + K + S)
color_regression <- color_regression_linear
color_differences_regression <- dE(predict(color_regression), master_colors %>% select("L", "a", "b")) %>% 
  data.frame(Difference = .)

# summary(color_regression)
```

With a multivariate multiple linear regression we find `r sum(color_differences_regression > 2)` colors that should be different on first glimpse. The median color difference is `r median(color_differences_regression$Difference)` (MAD = `r mad(color_differences_regression$Difference)`).

```{r, echo=FALSE, fig.width=8, fig.height=8}
color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% select(-Difference, -L, -a, -b) %>% bind_cols(
    color_differences_regression, predict(color_regression)
  ) %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference > 2),
            lwd = 1.5, width=0.95, height=0.95) +
  scale_color_manual(values = c("#FFFFFF00", "#CC3333")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = Color), size = 13) +
  scale_color_identity() +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 13
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
```

```{r, echo=FALSE}
degree <- 3
color_regression_poly <- lm(data = master_colors, cbind(L, a, b) ~ poly(C, degree) + poly(M, degree) + poly(Y, degree) + poly(K, degree) + poly(S, degree))
color_regression <- color_regression_poly
color_differences_regression <- dE(predict(color_regression), master_colors %>% select("L", "a", "b")) %>% 
  data.frame(Difference = .)

# summary(color_regression)
```

With a multivariate multiple polynomial regression with degree `r degree`  we find `r sum(color_differences_regression > 2)` colors that should be different on first glimpse. (38 colors with 2nd degree polynomials.) This are almost as many as with the linear regression. We even find some colors that have been matched better with linear regression. The median color difference is `r median(color_differences_regression$Difference)` (MAD = `r mad(color_differences_regression$Difference)`).

```{r, echo=FALSE, fig.width=8, fig.height=8}
color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% select(-Difference, -L, -a, -b) %>% bind_cols(
    color_differences_regression, predict(color_regression)
  ) %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference > 2),
            lwd = 1.5, width=0.95, height=0.95) +
  scale_color_manual(values = c("#FFFFFF00", "#CC3333")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = Color), size = 13) +
  scale_color_identity() +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 13
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
```

```{r, echo=FALSE}
color_regression_interaction_3terms <- lm(data = master_colors %>% select(C, M, Y, K, S, L, a, b), cbind(L, a, b) ~ .^3)
color_regression <- color_regression_interaction_3terms
color_differences_regression <- dE(predict(color_regression), master_colors %>% select("L", "a", "b")) %>% 
  data.frame(Difference = .)

# summary(color_regression)
```

With a multivariate multiple linear regression including interaction terms up to 3rd order we find `r sum(color_differences_regression > 2)` colors that should be different on first glimpse. (Five colors with only 2nd order interaction terms.) The median color difference is `r median(color_differences_regression$Difference)` (MAD = `r mad(color_differences_regression$Difference)`).

```{r, echo=FALSE, fig.width=8, fig.height=8}
color_differences %>% 
  filter(Sheet == 1, Row == 1, Column == 1) %>% select(-Difference, -L, -a, -b) %>% bind_cols(
    color_differences_regression, predict(color_regression)
  ) %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)) %>% 
  ggplot() +
  geom_tile(aes(fill = Difference, x = Ccol, y = Crow, color = Difference > 2),
            lwd = 1.5, width=0.95, height=0.95) +
  scale_color_manual(values = c("#FFFFFF00", "#CC3333")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(x = Ccol+0.175, y = Crow, color = Color), size = 13) +
  scale_color_identity() +
  scale_fill_continuous(limits = c(0, 10)) +
  geom_point(
    data = master_colors %>% rowwise() %>% mutate(Color = lab_to_rgb(L, a, b)),
    aes(x = Ccol-0.175, y = Crow, color = Color), size = 13
    ) +
  coord_fixed() +
  labs(caption = "Left circle: master color; Right circle: sample color") +
  xlab("Column") + ylab("Row") +
  theme(
    axis.text=element_text(size=12),
    axis.title=element_text(size=14,face="bold"),
    plot.caption = element_text(size=12, hjust = 0)
    )
```

**Train on part of data and assess performance on testdata (with bootstrap?)**

**Discuss parameter number and learning data: 25 vs 64 for 3rd order interaction (15 vs 64 in 2nd order interaction; 15 vs 64 in 3rd degree polinomial)**